Article Title

### 2013-12-20 plos.tex
rev27 (rev26 after responding to Kreso's changes/comments)

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects currently under way aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.  Correlations measured in such recordings are anticipated to reveal important aspects of the functional organization of neural circuits.  However, estimation and interpretation of large correlation matrices from finite recordings are challenging.  Estimation can be improved by regularization: biasing of the estimate toward a low-dimensional, less variable approximation.  The amount of improvement depends on how parsimoniously the reduced approximation captures the dominant dependencies in the data.  Therefore, the selection of the most efficient estimator is an empirical question that informs about the types of dominant dependencies governing the system.

In this study, we sought the most statistically efficient estimator of neural correlations in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 100 $\mu$m in diameter (150--350 cells) in mouse visual cortex.  We hypothesized that, in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of linear interactions between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.  Indeed, in cross-validation tests, the covariance matrix estimator that imposed this correlation structure outperformed estimators regularized toward other plausible low-dimensional approximations. Furthermore, we found consistent relationships between the organization of sparse interactions inferred by this estimator and the physical distances and preferred orientation differences of pairs of cells.  Consistent with previous synaptic connectivity studies, the density of positive interactions decreased rapidly with distance and with difference in preferred orientation whereas negative interactions were less selective.  To further corroborate physiological interpretation of the inferred functional structure, future experiments will augment this analysis with measured synaptic connectivity, cell types, and brain states.

\section*{Author Summary}
% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
Correlations of the activity of neurons have proven useful as descriptors of the functional organization of neural circuits with implications for stimulus coding and circuit architecture.  Estimation of correlation matrices can be improved by imposing some kind of structure on the estimate with greatest improvement attained when the imposed structure efficiently captures real dependencies in the data. Using fast 3D two-photon imaging of calcium signals, we recorded the activity of large and dense groups of cells in mouse visual cortex and evaluated the cross-validated performance of correlation matrix estimators that imposed different kinds of structure. The correlation structure of the estimator that proved most efficient comprised a sparse network of partial correlations between pairs of neurons combined with several common latent factors, with both components required for efficient estimation. Because of its superior benefit for estimation, we proposed that this inferred structure could prove more relevant for the description of functional connectivity than the usual correlation matrix in densely sampled neural recordings. As a first application of this approach, we analyzed how the inferred connectivity related to distances between cells and differences in their preferred orientations and found basic agreement with previous studies of synaptic connectivity.

\section*{Introduction}
\subsection*{Neural correlations}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been shown to have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Sompolinsky:2001,Wilke:2002,Averbeck:2006,Josic:2009,Berens:2011,Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits.  Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001,Arieli:1995,Chiu:2002,Kenet:2003,Kohn:2005,Cohen:2008,Cohen:2009,Ko:2011,Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.

@@

For example, one form of low-dimensional representation of the correlation matrix is its low-rank approximation. Low-rank approximations are particularly suitable for capturing shared fluctuations across the recorded population. Extracted through principal component or factor analysis, low-rank components of neural correlations have been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2009}.

@@

\subsection*{Conclusions and future work}
We proposed an alternative statistically verified method for estimating the functional connectivity in neural circuits at temporal scales attainable by calcium imaging.

>>>>

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.  Correlations in neural activity measured in such recordings can reveal important aspects of the functional organization of neural circuits.  However, estimating and interpreting large correlation matrices obtained from finite recordings is challenging.  Estimation can be improved by regularization: biasing of the usual unbiased estimate toward its low-dimensional approximation.  The amount of improvement is greatest when the approximation chosen as the target of regularization parsimoniously captures the dominant dependencies in the data.  Therefore, the construction of the most efficient estimator of neural correlation matrices is an empirical question that informs about the types of dominant dependencies governing the system.

In this study, we sought most statistically efficient estimators of neural correlations in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 100 $\mu$m in diameter (150--350 cells) in mouse visual cortex.  We hypothesized that in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of partial correlations between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.  Indeed, in cross-validation tests, the covariance matrix estimator based on such approximations outperformed other regularized estimators. Mirroring previously described patterns of synaptic connectivity, the density of positive partial correlations decreased rapidly with both physical distances and preferred orientation differences whereas negative partial correlations were less selective.  The inferred function connectivity obtained using these methods can provide hypotheses about synaptic architecture that can guide future experimental work.
\section*{Author Summary}
% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
Correlations in the activity of neurons have proven useful as descriptors of the functional organization of neural circuits with implications for stimulus coding and circuit architecture.  Estimation of correlation matrices can be improved by imposing some structure on the estimate. Greatest improvements are attained when the imposed structure efficiently captures real dependencies in the data. Using fast 3D two-photon imaging of calcium signals, we recorded the activity of large and dense groups of cells in mouse visual cortex and evaluated the performance of correlation matrix estimators that imposed different kinds of structure. The correlation structure of the estimator that led to greatest improvement comprised a sparse network of partial correlations between pairs of neurons combined with partial correlations with several common latent factors. Both components were necessary for efficient estimation. Since it leads to better estimates, we proposed that this sparse+latent structure provides a better picture of the  functional connectivity than that obtained from the usual sample correlation matrix in densely sampled neural recordings. As an application of this approach, we analyzed how the inferred connectivity related to distances between cells and differences in their preferred orientations and found basic agreement with previous studies of synaptic connectivity.

\section*{Introduction}
\subsection*{Neural correlations}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}~the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994,Abbott:1999,Sompolinsky:2001,Averbeck:2006,Josic:2009,Berens:2011,Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits.  This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001,Arieli:1995,Chiu:2002,Kenet:2003,Kohn:2005,Cohen:2008,Cohen:2009,Ko:2011,Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.

@@

For example, one form of low-dimensional representation of the correlation matrix is its low-rank approximation. Such approximations are particularly suitable for capturing shared fluctuations across the recorded population. Low-rank components can be  extracted using principal component or factor analysis. The resulting structure of  neural correlations has been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2009}.

@@
